\section{Roadmap}

Written up, many areas need a bit of polish:

\begin{itemize}
    \item Definition of SCDP
    \begin{itemize}
        \item Consequences (with/without side information)
        \item Causal theories \& prospects
    \end{itemize}
    \item Reduction of SCDP to SDP 
    \begin{itemize}
        \item known consequence kernel
        \item identifiable/loss-identifiable
        \item Reverse: SDP to SCDP
    \end{itemize}
    \item Defining causal prospects:
    \begin{itemize}
        \item Trivial (skepticism/helplessness)
        \item From classes of CBNs
        \item Universal independence with/without trivial action
    \end{itemize}
\end{itemize}


Overall comment: a lot of work on fitting existing theories into SCDP framework. It would be nice to have examples or results that demonstrate why the SCDP framework is useful (i.e. ones that are difficult to pose or prove in other frameworks).

To do/could be done.
\begin{itemize}
    \item Why it is necessary for consequences to be a kernel rather than JPD
    \item Examples:
    \begin{itemize}
        \item Where a non-physical model is preferred (even with infinite samples? this is tough)
        \item Conservation laws (\emph{where was it I've seen these before})
    \end{itemize}
    \item Lattice structure of causal prospects
    \begin{itemize}
        \item Any interesting inclusion results for nested Markov models, invariance sets, ADMGs or potential outcomes/Causal Bayes Nets?
    \end{itemize}
    \item Weakening of CBNs:
    \begin{itemize}
        \item Any legitimate examples of invariance but not CMC?
        \item What is a partially complete causal invariance network?
        \item Which identifiability results from CBNs are relevant to CINs?
        \item Conservative PC and faithfulness without CMC
    \end{itemize}
    \item Approximate loss-identifiability
    \begin{itemize}
        \item Define (straightforward from loss-identifiability)
        \item Interesting examples of approximate identifiability
    \end{itemize}
    \item ``Very soft interventions'' (i.e. limited information about distribution of target variables)
    \item Sequential generalisation (vectorise everything?)
    \item Potential outcomes:
    \begin{itemize}
        \item Relating ignorability to universal independence
        \item ``Affine-transformed loss identifiability''
        \item When can counterfactual performance measures be posed as losses? (Is there are reduction from identifiable counterfactual questions to ordinary SDPs?)
    \end{itemize}
    \item ``Model complexity''
    \begin{itemize}
        \item Generalisation of independence of conditionals to causal prospects
        \item Axiomatic approach to complexity measures
        \item Vague idea that the ``same'' measure should apply to $\mu$ and $\mathscr{T}$ - clarify this
    \end{itemize}
    \item Meta-learning connection: when is $\mathscr{T}(\mu)$ an easily learnable class? (Also connection to Finn's work)
\end{itemize}